Accelerated evolution of convective simulations

### 2018-02-21 tex/bvp_paper.tex
figure overhauls

\begin{abstract}
We present a method for coupling boundary value problems (BVPs) with initial value problems (IVPs)
in order to achieve thermally converged convective solutions on dynamical timescales, rather than the
long thermal timescale. We study this method in the context of \RB convection. 
We demonstrate that the solution reached by BVP and the
solution reached by a long thermal rundown of the IVP are similar, and demonstrate that this method works at a
large range of supercriticalities.  The BVP method is used to achieve converged solutions at high supercriticality ($10^8$),
and its extensions to more complex scenarios are discussed.
\end{abstract}
\maketitle

\section{Introduction}
\label{sec:intro}
Natural convection occurs in the presence of disparate timescales which
prohibit numericists from studying realistic models of natural systems.  For example,
flows in the Sun's convection zone range from a Mach number (Ma) of order unity
near the solar surface to order $10^{-5}$ deep in the solar interior.
Explicit methods which are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit must resolve the fastest motions (sound
waves and surface convection), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff. Even in the absence of the fast surface motions, the disparity between
the sound crossing time and the deep flows have made studies of low-Ma stellar
convection difficult. Traditionally, approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out,
have been used to study low-Ma flows \cite{brown&all2010, featherstone&hindman2016}.
More recently, advanced numerical techniques which use implicit or mixed
implicit-explicit timestepping mechanisms have made it feasible to study
convection at low Mach numbers \cite{viallet&all2011, viallet&all2013, viallet&all2016, lecoanet&all2014,
anders&brown2017, bordwell&all2018}, and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

While solutions to problem of disparate dynamical timescales have proven useful,
the difference between convective timescales and system relaxation timescales remains
a significant problem facing studies of convection. 
As modern simulations aim to model natural convection
by increasing into the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale on which the system relaxes grows with respect
to the dynamical timescale, which stays roughly constant across all Ra
\cite{anders&brown2017}.
Solar convection is, once again, a prime example of this phenomenon.
Dynamical timescales in the Solar convective zone are relatively short 
(10 min overturn at solar surface, one month solar
rotation rate) compared to the Kelvin-Helmholtz timescale of
$3 \cdot 10^7$ years \cite{stix2003}.  
Furthermore, as dynamical and diffusive timescales separate, 
simulations become more turbulent. More turbulent motions 
require finer grid meshes and smaller timesteps
to capture advective dynamics. Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
while the number of overturn times required for systems to reach thermal equilibration
grows.

The vast difference between convective and diffusive timescales has long plagued
numericists studying convection, and a plethora of approaches has been employed to
study thermally converged solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``Bootstrapping,'' or the process of using the flow
fields in a converged solution at lower Ra as initial conditions for a simulation at a higher
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  Bootstrapped solutions are succeptible to hysteresis
effects, in which large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
In moderate-Ra simulations, a commonly-used tactic is to use 
a simple model of the full convective state as initial conditions.  
For example, past studies have used a linear eigenvalue solve to set the initial
convective state \cite{hurlburt&all1984} or used an axisymmetric solution 
as initial conditions for convection in a 3D cylinder \cite{verzicco&camussi1997}. 
In certain systems, 
the approximate state of the evolved solution can be estimated, and there an
appropriate set of initial conditions can either be solved for analytically
\cite{couston&all2017} or by using knowledge of Mixing Length Theory or other convective
theories to adjust the initial profile towards the proper adiabatic state \cite{brandenburg&all2005}.

@@

In this work, we study a method of achieving accelerated evolution of a
convective simulation through adjusting the thermodynamic state using information 
from resolved, convective dynamics. We couple measurements of the (non-converged)
convective solution with knowledge about energy balances in the eventual solution
to self-consistently adjust the mean thermodynamic profile towards. 
While such a method has been used previously \cite{hurlburt&all1986}, 
we find no explanation in the current literature of the steps involved in employing
this method, nor any study into the accuracy of such a method.
In section \ref{sec:experiment}, we describe our convective simulations, our
numerical methods, and our method for achieving accelerated evolution. In
section \ref{sec:results}, we compare solutions reached through the accelerated evolution
method to those that have been evolved through a full diffusive timescale, and we
examine select simulations at very high Ra which have achieved accelerated evolution. Finally,
in section \ref{sec:discussion}, we discuss extensions of the methods presented
here to stratified, compressible systems.

\section{Experiment}
\label{sec:experiment}
We study incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that our fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$, except on the
term where the constant gravitational acceleration, $\bm{g} = - g\hat{z}$, acts in the vertical momentum equation, 
where $\rho = \rho_0(1  - \alpha T_1)$.  
Under these constraints, the equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$ is the velocity, 
$T = T_0 + T_1$ are the initial and fluctuating components of temperature, 
and $P$ is the pressure.
We non-dimensionalize these equations such that the dimensionless unit of
length is the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$).
by these choices, one time unit is a freefall time ($L_z/v_{ff}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv P / \rho_0 + \phi + |\bm{u}|^2 / 2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. As $P$ is a
Langrange multiplier under the Oberbeck-Boussinesq approximation, $\varpi$
can be treated straightforwardly as a linear variable. 
In non-dimensional form, Eqns. \ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
are
\begin{gather}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega}
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1.
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{P}$ and $\mathcal{R}$ 
are set by the Rayleigh and Prandtl numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr$ = 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.

We study 2D and 3D convection in which the domain is a cartesian box, 
whose dimensionless vertical extent is $z \in [-1/2, 1/2]$, 
and which is horizontally period with an extent of $x, y \in [0, \Gamma]$ where $\Gamma = 2$ is the aspect
ratio. 
In the 2D cases, we set $v = \partial_y = 0$.
We specify no-slip, impenetrable boundary conditions at both the top and
bottom boundary and we use mixed thermal boundary conditions, such that
\begin{equation}
u = v = w \, \, \text{at}\,\,z = \pm 1/2, \qquad T_1 = 0 \,\,\text{at}\,\, z=+1/2, \qquad
\frac{\partial T_1}{\partial z} = 0\,\,\text{at}\,\,z=-1/2.
\label{eqn:bcs}
\end{equation}
For this choice of boundary conditions, the critical value of Ra at which
the onset of convection occurs is Ra$_{\text{crit}} = 1295.78$, and the
supercriticality of a run is defined as $S \equiv \text{Ra}/\text{Ra}_{\text{crit}}$.
Studies of convection which aim to model
astrophysical systems such as stars often employ mixed thermal
boundary conditions \cite{hurlburt&all1984, cattaneo&all1991, korre&all2017},
as we do here; however, our choice of thermal boundary conditions here
reflects the fact that the conditions in Eqn. \ref{eqn:bcs} are the simplest
to implement in the process of accelerated evolution (see section \ref{subsection:ae})
we study here.

@@

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\text{cond}} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective flux,
F$_{\text{conv}} = \angles{w(T_0 + T_1)}_{x,y}$, are the basis of the AE method. We measure
both of these quantities early in a simulation, as in Fig. \ref{fig:time_trace}c.
At these early stages in the simulation, these flux profiles are highly asymmetric,
with much more flux leaving through the upper boundary than entering through the
lowe boundary as the atmosphere approaches the proper isotherm selected by the
fixed temperature upper boundary.
However, by calculating the total flux,
F$_{\text{tot}} = $F$_{\text{conv}} +$ F$_{\text{cond}}$, and then calculating the profiles
\begin{equation}
f_{\text{conv}}(z) = \frac{F_{\text{conv}}}{F_{\text{tot}}},\qquad
f_{\text{cond}}(z) = \frac{F_{\text{cond}}}{F_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the early thermal boundary layers are roughly the proper
width.  Where $f_{\text{conv}} = 1$, convection dominates all transport, and where
$F_{\text{cond}} = 1$, conduction dominates all transport (in the boundary
layers). Under this assumption, the proper evolved atmospheric flux profiles
are F$_{\text{conv, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{conv}}$
and F$_{\text{cond, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{cond}}$,
where F$_{\text{bot}} = \mathcal{P}$ is the amount of flux entering the
bottom of the atmosphere.

@@

The AE method is thus simple: we construct F$_{\text{conv, ev}}$ as described above.
Then we calculate a profile, 
$\xi(z) = \text{F}_{\text{conv, ev}}/\text{F}_{\text{conv}}$, which is the amount that the
flux in the system needs to be reduced by, as a function of height.
We multiply the velocities
and the thermal fluctuations, $T - \angles{T}$, by $\sqrt{\xi}$, such that the product of all fluctuations
(which carry the convective flux) are diminished by a factor of $\xi$.  We solve
Eqns. (\ref{eqn:bouss_BVP_momentum}-\ref{eqn:bouss_BVP_energy}) with F$_{\text{conv, ev}}$
and $\xi\cdot\angles{\bm{u}\times\bm{\omega}}_{x,y}$ for the mean thermodynamic state,
and then continue evolving in time.  This adjustment of the mean profile and the
diminishing of velocities and fluctuations is the AE method, and it can generally
be applied tens of buoyancy times after the peak of convective transient.

@@

As the AE method primarily serves to adjust the thermodynamic structure of the
solution, we compare the temperature profile attained by AE and SE in 
Fig. \ref{fig:temp_comparison}.  We see that the boundary layer length scale is 
nearly identical in the AE and SE solution (Fig. \ref{fig:temp_comparison}a), but that
the mean temperature in the interior differs by about 0.5\% on average
(Fig. \ref{fig:temp_comparison}b). The probability distribution function of point-by-point
temperature values are compared in Fig. \ref{fig:temp_comparison}c.  We construct
this PDF by interpolating our temperature field onto a regular grid, determing the
frequency distribution of all $T$ values, and then properly normalizing such that
the integral of the PDF is unity.  In addition to the
difference in mode which can be seen in the mean vertical profile, the SE case
has a broader spread of temperature values.  This can likely be attributed to the
slight overstabilization of the temperature profile obtained by the AE method
(as can be seen in Fig. \ref{fig:time_trace}a\&b).  One means of comparing two
probability distributions to determine if they are drawn from the same underlying
sample is through the use of a Kolmogorov-Smirnov (KS) \cite{wall&jenkins2012}.
In general, a KS test must be conducted on independent, uncorrelated data, which
poorly describes the point-by-point values of flow in a fluid simulation. Thus,
we will merely use the KS statistic, the maximum difference
between the cumulative distribution functions (CDFs) of the two sample distributions,
as a numerical method of directly comparing the two PDFs.  For the distributions
shown in Fig. \ref{fig:temp_comparison}c,
we find a KS statistic of 0.113 near the modes, which is large, and implies that
these two distributions are different.

@@

In addition to comparing the thermodynamic state achieved by the SE and AE methods,
we examine the velocities found in the evolved states.
We compute PDFs in the same manner as in Fig. \ref{fig:temp_comparison}c for the
vertical velocity (Fig. \ref{fig:pdf_comparison}a), horizontal velocity (Fig. \ref{fig:pdf_comparison}b),
and the nonlinear convective flux (Fig. \ref{fig:pdf_comparison}c). We report KS statistics
of 0.0577, 0.858, and 0.0347, respectively.  Perhaps unsurpisingly, the nonlinear
convective transport between the SE and AE methods are very similar, as is captured
in a volume-averaged sense in the Nu measurements of Fig. \ref{fig:parameter_space_comparison}a\&d).
In general, the velocity boundary conditions (Eqn. \ref{eqn:bcs}) more strongly dominate the flow
fields for the SE cases than the AE cases in Fig. \ref{fig:pdf_comparison}a\&c. 
This suggests that the velocity boundary layers are not yet fully formed in the
AE solutions. However, aside from this difference, the two solutions show a similar
distribution of realized velocity values.

@@

Despite differences between the SE and AE solutions for the case studied in 
Figs. \ref{fig:time_trace}, \ref{fig:temp_comparison}, \& \ref{fig:pdf_comparison},
the AE method is still extremely powerful.  The first application of the AE method
($t \approx 70$ in Fig. \ref{time_trace}b) immediately increases the 
average time step of our solver by a factor of 2-3. At higher supercriticality
($S = 10^7$), the AE solve immediately boosts the timestep by nearly a factor of 4.
Thus, not only does this method evolve the solution into nearly the correct state, 
but further time evolution (either to achieve precisely the correct thermodynamic
state or to take measurements of fluid quantities) happens more efficiently.

\section{Extensions of the AE method}
\label{section:extensions}
Here we have used \RB convection as a test case for the AE method, but we argue that
the true power of this method is in its extensions to more complicated studies.
Such extensions are beyond the scope of this work, but we will briefly describe
avenues in which extensions to the AE method will be useful.
For example, in studies of stratified convection where the thermal diffusivity
is inversely proportional to the density \cite{anders&brown2017}, AE solutions
will allow for the careful study of a broader range of parameter space. 
Further, in studies where stable and convecting regions are paired
\cite{hurlburt&all1986, couston&all2017}, the stable layers cannot rely on convective
motions to restratify them and instead evolve on the order of the slow thermal
timescale. In
fully-convective domains, such as those studied here, the atmosphere evolves
much more rapidly than the thermal diffusion time (see e.g., Fig. \ref{fig:time_trace}a).

@@

\section{Conclusion}
\label{sec:discussion}
Here we have described a method for converging convective simulations using
a method of Accelerated Evolution (AE), and compared it to Standard
Evolution (SE) achieved by running simulations for a thermal timescale.
We have shown that the AE method produces dynamics that are similar to SE
dynamics on much shorter timescales, and have discussed differences that arise
between the two methods.

>>>>

\begin{abstract}
We present a method for coupling boundary value problems with initial value problems 
in order to achieve Accelerated Evolution (AE) of convective solutions on
dynamical timescales, rather than the long thermal timescale. 
We study this method in the context of \RB convection. 
We demonstrate that the solution reached by AE and Standard Evolution (SE) are
similar, and that this method works at a
large range of supercriticalities.  The AE method is used to achieve converged 
solutions at high supercriticality ($10^7$), and its extensions to more complex 
systems are briefly discussed.
\end{abstract}
\maketitle

\section{Introduction}
\label{sec:intro}
Natural convection occurs in the presence of disparate timescales which
prohibit numericists from studying realistic models of natural systems.  For example,
flows in the convection zone of stars like the Sun are often at low Mach number
(Ma) deep in the interior.
Explicit methods which are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit must resolve the fastest motions (sound
waves), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff, and the disparity between
the sound crossing time and the convective flows have made studies of low-Ma stellar
convection difficult. Traditionally, approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out,
have been used to study low-Ma flows \cite{brown&all2010, featherstone&hindman2016}.
More recently, advanced numerical techniques which use implicit or mixed
implicit-explicit timestepping mechanisms have made it feasible to study
convection at low Mach numbers \cite{viallet&all2011, viallet&all2013, viallet&all2016, lecoanet&all2014,
anders&brown2017, bordwell&all2018}, and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

While solutions to problem of disparate \emph{dynamical} timescales have proven useful,
the difference between convective timescales and system relaxation timescales remains
a significant problem facing studies of convection. 
As modern simulations aim to model natural convection
by increasing into the high-Rayleigh-number (Ra) regime,
the ratio of the thermal diffusion timescale to dynamical timescales
grows \cite{anders&brown2017}.
Solar convection is a prime example of this phenomenon.
Dynamical timescales in the solar convective zone are relatively short 
(10 min overturn at solar surface, one month solar
rotation rate) compared to the Kelvin-Helmholtz timescale of
$3 \cdot 10^7$ years \cite{stix2003}.  
Furthermore, as dynamical and diffusive timescales separate, 
simulations become more turbulent. Increasingly turbulent motions 
require finer grid meshes and smaller timesteps
to capture advective dynamics. Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
while the number of overturn times required for systems to reach thermal equilibration
grows.

The vast difference between convective and diffusive timescales has long plagued
numericists studying convection, and a plethora of approaches has been employed to
study thermally converged solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``Bootstrapping,'' or the process of using the flow
fields in a converged solution at lower Ra as initial conditions for a simulation at a higher
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  Bootstrapped solutions are succeptible to hysteresis
effects, in which large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
Another commonly-used tactic in
moderate-Ra simulations is to use 
a simple model of the full convective state as initial conditions.  
For example, past studies have used a linear eigenvalue solve to set the initial
convective state \cite{hurlburt&all1984} or used an axisymmetric solution 
as initial conditions for convection in a 3D cylinder \cite{verzicco&camussi1997}. 
In certain systems, 
the approximate state of the evolved solution can be estimated. There, an
appropriate set of initial conditions can either be solved for analytically
\cite{couston&all2017} or by using knowledge of Mixing Length Theory or other convective
theories to adjust the initial profile towards the proper adiabatic state \cite{brandenburg&all2005}.

@@

In this work, we study a method of achieving accelerated evolution of a
convective simulation through adjusting the thermodynamic state using information 
from resolved, convective dynamics. We couple measurements of the (non-converged)
convective solution with knowledge about energy balances in the eventual solution
to self-consistently adjust the mean thermodynamic profile towards its final state. 
While such a method has been used previously \cite{hurlburt&all1986}, 
we find no explanation in the current literature of the steps involved in employing
this method, nor any study into the accuracy of such a method.
In section \ref{sec:experiment}, we describe our convective simulations, our
numerical methods, and our method for achieving accelerated evolution. In
section \ref{sec:results}, we compare solutions reached through the accelerated evolution
method to those that have been evolved through a full diffusive timescale, and we
examine select simulations at very high Ra which have achieved accelerated evolution. Finally,
in section \ref{sec:extensions}, we discuss extensions of the methods presented here,
and we offer concluding remarks in section \ref{sec:conclusion}.

\section{Experiment}
\label{sec:experiment}
We study incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that our fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$, except on the
term where the constant gravitational acceleration, $\bm{g} = - g\hat{z}$, acts in the vertical momentum equation, 
where $\rho = \rho_0(1  - \alpha T_1)$.  
Under these constraints, the equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$ is the velocity, 
$T = T_0 + T_1$ are the initial and fluctuating components of temperature, 
and $P$ is the pressure.
We non-dimensionalize these equations such that the dimensionless unit of
length is the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$).
by these choices, one time unit is a freefall time ($L_z/v_{ff}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv P / \rho_0 + \phi + |\bm{u}|^2 / 2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. As $P$ is a
Langrange multiplier under the Oberbeck-Boussinesq approximation, $\varpi$
can be treated straightforwardly as a linear variable. 
In non-dimensional form, Eqns. \ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{gather}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1.
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{P}$ and $\mathcal{R}$ 
are set by the Rayleigh and Prandtl numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr$ = 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.

We study 2D and 3D convection in which the domain is a cartesian box, 
whose dimensionless vertical extent is $z \in [-1/2, 1/2]$, 
and which is horizontally periodic with an extent of $x, y \in [0, \Gamma]$ where $\Gamma = 2$ is the aspect
ratio. 
In the 2D cases, we set $v = \partial_y = 0$.
We specify no-slip, impenetrable boundary conditions at both the top and
bottom boundary and we use mixed thermal boundary conditions, such that
\begin{equation}
u = v = w = 0 \, \, \text{at}\,\,z = \pm 1/2, \qquad T_1 = 0 \,\,\text{at}\,\, z=+1/2, \qquad
\frac{\partial T_1}{\partial z} = 0\,\,\text{at}\,\,z=-1/2.
\label{eqn:bcs}
\end{equation}
For this choice of boundary conditions, the critical value of Ra at which
the onset of convection occurs is Ra$_{\text{crit}} = 1295.78$, and the
supercriticality of a run is defined as $S \equiv \text{Ra}/\text{Ra}_{\text{crit}}$.
Studies of convection which aim to model
astrophysical systems such as stars often employ mixed thermal
boundary conditions \cite{hurlburt&all1984, cattaneo&all1991, korre&all2017},
as we do here; however, our choice of thermal boundary conditions here
reflects the fact that the conditions in Eqn. \ref{eqn:bcs} are the simplest
to implement in the process of accelerated evolution (see section \ref{subsection:ae})
we study here.

@@

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\text{cond}} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective flux,
F$_{\text{conv}} = \angles{w(T_0 + T_1)}_{x,y}$, where $\angles{}_{x,y}$ represent
a horizontal average, are the basis of the AE method. We measure
both of these quantities early in a simulation, as in Fig. \ref{fig:time_trace}c.
At these early stages in the simulation, these flux profiles are highly asymmetric,
with much more flux leaving through the upper boundary than entering through the
lower boundary as the atmosphere approaches the proper isotherm selected by the
fixed temperature upper boundary condition.
However, by calculating the total flux,
F$_{\text{tot}} = $F$_{\text{conv}} +$ F$_{\text{cond}}$, and then calculating the profiles
\begin{equation}
f_{\text{conv}}(z) = \frac{F_{\text{conv}}}{F_{\text{tot}}},\qquad
f_{\text{cond}}(z) = \frac{F_{\text{cond}}}{F_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
we remove the asymmetry and examine which parts of the atmosphere are dominated
by conduction, and which are not.
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the early thermal boundary layers are roughly the proper
width.  Where $f_{\text{conv}} = 1$, convection dominates all transport, and where
$f_{\text{cond}} = 1$, conduction dominates all transport (in the boundary
layers). Under this assumption, the proper evolved atmospheric flux profiles
are F$_{\text{conv, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{conv}}$
and F$_{\text{cond, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{cond}}$,
where F$_{\text{bot}} = \mathcal{P}$ is the amount of flux entering the
bottom of the atmosphere.

@@

The AE method is thus simple: we construct F$_{\text{conv, ev}}$ as described above.
Then we calculate a profile, 
$\xi(z) = \text{F}_{\text{conv, ev}}/\text{F}_{\text{conv}}$, which is the amount that the
flux in the system needs to be reduced by, as a function of height.
We multiply the velocities
and the thermal fluctuations, $T - \angles{T}$, by $\sqrt{\xi}$, such that the product of all fluctuations
(which carry the convective flux) are diminished by a factor of $\xi$.  We solve
Eqns. (\ref{eqn:bouss_BVP_momentum}-\ref{eqn:bouss_BVP_energy}) with F$_{\text{conv, ev}}$
and $\xi\cdot\angles{\bm{u}\times\bm{\omega}}_{x,y}$ for the mean thermodynamic state,
and then continue evolving in time.  This adjustment of the mean profile and the
diminishing of velocities and fluctuations is the AE method, and it can generally
be applied tens of buoyancy times after the peak of convective transient.
For specifics on the precise implementation of the AE method, we refer
the reader to appendix \ref{appendix:recipe}.

@@

As the AE method primarily serves to adjust the thermodynamic structure of the
solution, we compare the temperature profile attained by AE and SE in 
Fig. \ref{fig:temp_comparison}.  We see that the boundary layer length scale is 
nearly identical in the AE and SE solution (Fig. \ref{fig:temp_comparison}a), but that
the mean temperature in the interior differs by about 0.5\% on average
(Fig. \ref{fig:temp_comparison}b). The probability distribution function of point-by-point
temperature values are compared in Fig. \ref{fig:temp_comparison}c.  We construct
this PDF by interpolating our temperature field onto a regular grid, determing the
frequency distribution of all $T$ values, and then properly normalizing such that
the integral of the PDF is unity.  In addition to the
difference in mode which can be seen in the mean vertical profile, the SE case
has a broader spread of temperature values.  This can likely be attributed to the
slight overstabilization of the temperature profile obtained by the AE method
(as can be seen in Fig. \ref{fig:time_trace}a\&b).  One means of comparing two
probability distributions to determine if they are drawn from the same underlying
sample is through the use of a Kolmogorov-Smirnov (KS) \cite{wall&jenkins2012}.
In general, a KS test must be conducted on independent, uncorrelated data, which
poorly describes the point-by-point values of flow in a fluid simulation. Thus,
we will merely use the KS statistic, the maximum difference
between the cumulative distribution functions (CDFs) of the two sample distributions,
as a numerical method of directly comparing the two PDFs.  For the distributions
shown in Fig. \ref{fig:temp_comparison}c,
we find a KS statistic of 0.185 near the modes, which is large, and implies that
these two distributions are different.

@@

In addition to comparing the thermodynamic state achieved by the SE and AE methods,
we examine the velocities found in the evolved states.
We compute PDFs in the same manner as in Fig. \ref{fig:temp_comparison}c for the
vertical velocity (Fig. \ref{fig:pdf_comparison}a), horizontal velocity (Fig. \ref{fig:pdf_comparison}b),
and the nonlinear convective flux (Fig. \ref{fig:pdf_comparison}c). We report KS statistics
of 0.00615, 0.0349, and 0.0204, respectively.  Perhaps unsurpisingly, the nonlinear
convective transport between the SE and AE methods are very similar, as is captured
in a volume-averaged sense in the Nu measurements of Fig. \ref{fig:parameter_space_comparison}a\&d).
In general, the velocity boundary conditions (Eqn. \ref{eqn:bcs}) more strongly dominate the flow
fields for the SE cases than the AE cases in Fig. \ref{fig:pdf_comparison}a\&c. 
This suggests that the velocity boundary layers are not yet fully formed in the
AE solutions. However, aside from this difference, the two solutions show a similar
distribution of realized velocity values.

@@

Despite differences between the SE and AE solutions for the case studied in 
Figs. \ref{fig:time_trace}, \ref{fig:temp_comparison}, \& \ref{fig:pdf_comparison},
the AE method is still extremely powerful.  The first application of the AE method
($t \approx 70$ in Fig. \ref{fig:time_trace}b) immediately increases the 
average time step of our solver by a factor of 2-3. At higher supercriticality
($S = 10^7$), the AE solve immediately boosts the timestep by nearly a factor of 4.
Thus, not only does this method evolve the solution into nearly the correct state, 
but further time evolution (either to achieve precisely the correct thermodynamic
state or to take measurements of fluid quantities) happens more efficiently.

\section{Extensions of the AE method}
\label{sec:extensions}
Here we have used \RB convection as a test case for the AE method, but we argue that
the true power of this method is in its extensions to more complicated studies.
Such extensions are beyond the scope of this work, but we will briefly describe
avenues in which extensions to the AE method will be useful.
For example, in studies of stratified convection where the thermal diffusivity
is inversely proportional to the density \cite{anders&brown2017}, AE solutions
will allow for the careful study of a broader range of parameter space. 
Further, in studies where stable and convecting regions are paired
\cite{hurlburt&all1986, couston&all2017}, the stable layers cannot rely on convective
motions to restratify them and instead evolve on the order of the slow thermal
timescale. In
fully-convective domains, such as those studied here, the atmosphere evolves
much more rapidly than the thermal diffusion time (see e.g., Fig. \ref{fig:time_trace}a).

@@

\section{Conclusion}
\label{sec:conclusion}
Here we have described a method for converging convective simulations using
a method of Accelerated Evolution (AE), and compared it to Standard
Evolution (SE) achieved by running simulations for a thermal timescale.
We have shown that the AE method produces dynamics that are similar to SE
dynamics on much shorter timescales, and have discussed differences that arise
between the two methods.
